---
title: "index"
output: html_document
date: "2023-02-23"
---
# Load Data
```{r, load-packages}
library("tidyverse")
library("languageR")
```

# Data Set 
```{r, data}
beginningReaders <- beginningReaders
```

# Descriptive Data
```{r, descriptives}
beginningReaders %>%
  group_by(Word) %>% 
  summarise(Mean = mean(LogRT), 
            SD = sd(LogRT))
```

# Linear Method Correlation
```{r, bivariate-regression}
LogRT_Frequency <- lm(LogRT ~ LogFrequency, data = beginningReaders)
summary(LogRT_Frequency)
```

# Plot
```{r, LogRT-Frequency-Graph}
beginningReaders %>%
  ggplot(aes(x = LogFrequency, y = LogRT)) + 
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "Word Frequency", y = "RT", title = "RT by Word Frequency")

```

# Observation
The dataset selected was beginningReaders. To tidy the data, the dataset's words were grouped by the mean and standard deviation. The goal of a tidy data set is to have every column as a variable, every row as an observation, and ever cells as a single value. 

The linear model tests the relationship between the reaction time and the frequency of the words. It was then plotted as the x variable being word frequency and the y variable being the reaction time. The blue line indicated the correlation and R was asked to look for a linear model correlation. The flatter the line appears, the weaker the correlation. 



