---
title: "Programming assignment 2"  
author: "Meritxell Feliu Ribas"  
date: "Last update: `r Sys.Date()`"  
output:  
  html_document:  
    highlight: kate  
    keep_md: yes  
    theme: "united"
---

```{r, load-packages}
library("tidyverse")
library("languageR")
```

```{r, load data}
beginningReaders = beginningReaders
danish -> danish
dativeSimplified -> dativeSimplified
english -> english
spanishFunctionWords -> spanishFunctionWords
```

# Bivariate Scatterplot
```{r, bivariate-scatterplot}
beginningReaders %>% 
  ggplot() + 
  aes(x = LogFrequency, y = LogRT) + 
  geom_point() +
  labs(y = "Reaction Time", x = "Word Frequency",
       title = "Reaction Times by Word Frequency")
```

# Boxplot with different fill colors
```{r, boxplot-with-different-fill-colors}
danish %>% 
  ggplot() + 
  aes(x = Affix, y = LogRT, fill = Sex) + 
  geom_boxplot() +
  labs(x = "Affix", y = "Reaction Time",
       title = "Reaction Times by Affix and Sex")
```

# Scatterplot with Stat Summary and Facet
```{r, plot-of-my-choice-with-statsummary-facet}
english %>% 
  ggplot() + 
  aes(x = WordCategory, y = RTnaming) + 
  facet_grid(. ~ AgeSubject) + 
  stat_summary(fun.data = mean_se, geom = "pointrange", pch = 19) + 
  scale_y_continuous(breaks = seq(0, 10, 2),
                     limits = c(0, 10)) +
  labs(x = "Word Category", y = "Naming Reaction Time",
       title = "Naming Reaction Times by Word Category and Age")
```

# Comments

I selected the variables and datasets that I wanted to use and I did all the graphs. However, for the last one, I realized that the scale for the y axis was difficult to interpret (it started in 6) and I asked Patrick for help. He helped me to adjust the y axis. I hope it is fine.


