---
title: "Programming assignment 2"  
author: "Eva Corregidor"  
date: "Last update: `r Sys.time()`"  
output:  
  html_document:  
    highlight: kate  
    keep_md: yes  
    theme: united
---

install.packages("languageR")
library("languageR")

 # Familiarizing myself with datatets

data(beginningReaders)
summary(beginningReaders)
?beginningReaders

data(danish)
?danish

data(dativeSimplified)
?dativeSimplified

data(english)
?english

data(spanishFunctionWords)
?spanishFunctionWords

library("ggplot2")
library("ggthemes")

 # Create a bivariate_scatterplot

```{r}

data(english)
ggplot(
  data = english,
  aes(x = Familiarity, y = WrittenFrequency) 
)+
  geom_point(aes(color = LengthInLetters, shape = WordCategory)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Written frequency and familiarity (English)",
    subtitle = "Written frequency and subjective familiarity ratings",
    x = "Familarity (0-7)", y = "Written Frequency (0-12)",
    color = "Word Length, i.e number of letters", shape = "Word Category"
)
```

 # Create a boxplot with different fill colors

```{r}
ggplot(danish, aes(x = PrevError, y = LogRT, color = Sex, fill = Sex)
) +
  geom_boxplot() +
    labs(
    title = "Auditory lexical decision latencies for Danish complex words",
    subtitle = "Relation between the preciding trial and the RTs",
    x = "Preciding trial: correct lexical decision or error", y = "Response time (ms)",
    color = "Sex")
```

 # Create a plot of my choice (density_plot) that includes a stat_summary and a facet

```{r}

my_min <- min(dativeSimplified$LengthOfTheme)
my_max <- max(dativeSimplified$LengthOfTheme)

ggplot(dativeSimplified, aes(x = LengthOfTheme, fill = RealizationOfRec, color = RealizationOfRec)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~AnimacyOfRec) +
  scale_fill_viridis_d(begin = 0.2, end = 0.8) + 
  scale_color_viridis_d(guide = "none", begin = 0.2, end = 0.8) + 
  geom_vline(xintercept = my_min) +
  geom_vline(xintercept = my_max) +
  labs(
    title = "Dative Alternation",
    subtitle = "Distribution of number of words in the realization of the dative as NP or PP based on the animacy of the recipient",
    x = "Number of words", y = "Density",
    fill = "Dative Realization, dot = mean"
    )
```

 # (actual_plot_stat_summary) Create a plot of my choice (density_plot) that includes a stat_summary and a facet`
 
```{r}
ggplot(dativeSimplified, aes(x = AnimacyOfTheme, y = LengthOfTheme, color = RealizationOfRec)) +
  facet_wrap(~RealizationOfRec) +
  geom_boxplot() +
  stat_summary(fun.data = "mean_sdl") +
  labs(
    title = "Dative Simplified",
    subtitle = "Distribution of the animacy of the theme in relation to the number of words comprising the theme, organized by the realization of the dative (NP or PP)",
    x = "Animacy of theme", y = "Length of theme",
    color = "Realization of rec")
```
 