---
title: "Programming assignment 2"
author: "Ziling Zhu"
date: "Last update: `r Sys.Date()`"
output:
  github_document:
  highlight: kate
  keep_md: yes
  theme: united
---
## Setup
```{r}
library(languageR)
library(tidyverse)
```

## Scatterplot
Based on the data in the package **english**, the following scatterplot shows that the more frequently a word is written, the more familiar English users rate it to be.

```{r, echo = TRUE}
english %>% 
  ggplot() +
  aes(x = WrittenFrequency, y = Familiarity) +
  geom_point(alpha = 0.05) +
  geom_smooth(color = "navy", size = 0.5) +
  labs(x = "Written Frequency", y = "Familiarity",
       title = "Familiary as a function of Written Frequency") +
  theme_minimal()
```

## Boxplot
Based on the data in the package **dativeSimplified**, the following boxplot shows that animate themes have more words than inanimate themes.

```{r, echo = TRUE}
dativeSimplified %>% 
  ggplot() +
  aes(x = AnimacyOfTheme, y = LengthOfTheme, fill = AnimacyOfTheme) +
  geom_boxplot(width = 0.5, alpha = 0.5, show.legend = FALSE) +
  labs(x = "Animacy of Theme", y = "Length of Theme",
       title = "Length of Theme as a function of Animacy of Theme") +
  theme_minimal()
```

## Plot of my choice
Based on the data in the package **danish**, the following violin plot shows the gender variations in the (log-transformed) response time for each affix (not sure if it's linguistically meaningful).

```{r, echo = TRUE}
danish %>%
  ggplot() + 
  aes(x = Affix, y = LogRT, fill = Sex) +
  geom_violin(alpha = 0.5) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange",
               alpha = 0.5, color = "navy") +
  facet_grid(Sex ~ .) +
  labs(x = "Affix", y = "LogRT",
       title = "Response Time as a function of Affix grouped by Gender",
       fill = "Gender") +
  theme_minimal()
```