---
title: "Index"
output: html_document
author: "Meritxell Feliu Ribas"
date: "2023-02-23"
---

# Load Packages
```{r, load-packages}
library("tidyverse")
library("languageR")
```

# Load Data
```{r, load-data}
beginningReaders = beginningReaders
```

# Descriptive Statistics
```{r, summary-descriptive-statistics}
beginningReaders %>%
  group_by(Word) %>% 
  summarize(
    mean = mean(LogRT),
    sd = sd(LogRT),
    min = min(LogRT),
    max = max(LogRT)
    )
```

# Linear Model Correlation
```{r, correlation}
rt_freq <- lm(LogRT ~ LogFrequency, data = beginningReaders)
summary(rt_freq)
```

# Plot with regression line
```{r, plot}
beginningReaders %>% 
  ggplot() + 
  aes(x = LogFrequency, y = LogRT) + 
  geom_point() +
  geom_smooth(method = lm) +
  labs(y = "Reaction Time", x = "Word Frequency",
       title = "Reaction Times by Word Frequency")
```

# General Observations
The data set chosen for this analysis is BeginningReaders, which provides information about visual lexical decision with 8-year-old Dutch children. We have analyzed the effect of word frequency (independent variable, x) on reaction times (dependent variable, y) for a total of 184 words. As we can see from the R-squared value calculated from the model and the plot with the regression line, we find a weak negative correlation between RT and word frequency. 

# Comment
I worked in this assignment with Stephanie and Patrick, which helped me understand better what we did in class the other day. However, I wrote the general observations alone. 
