---
title: "Programming assignment 2"
author: "Laura Ramirez Polo"
date: "Last update:`r Sys.Date()`"
output: 
  html_document:
    highlight: kate 
    keep_md: yes
    theme: "united"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, 'load-data', echo = TRUE, warning = FALSE}
library("tidyverse")
library("languageR")
library("ggplot2")
library("dplyr")
```

## A bivariate scatter plot

```{r}
english %>%
  ggplot() + 
  aes(x = WrittenFrequency, y = WrittenSpokenFrequencyRatio) +
  geom_point()
```

## A boxplot with different fill colors
```{r}
spanishMeta %>%
  ggplot () +
  aes(x = FullName, y = Nwords) + 
  geom_boxplot(aes(colour=YearOfBirth))
```

## A plot of your choice that includes a stat_summary and a facet.

```{r}
beginningReaders %>%
  ggplot() +
  aes (x = ProportionOfErrors, y = PC1) +
  facet_grid(. ~ OrthLength) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.4) +
  stat_summary(fun.data = mean_se, geom = "point", pch = 21)
```

## Trying other plots
#Word seems to have a lot of observations, so it looks very cluttered. I wonder how this can be solved, if Word, a string, can be grouped with stat_summary? I tried with what we learned in class with "select" and "filter" to just show words starting with A, but it does not work. I will keep trying!

```{r}
beginningReaders %>%
  select(Word,ProportionOfErrors) %>% 
  filter(Word =="^A") %>% 
  ggplot () +
  aes(x = Word, y =ProportionOfErrors) + 
  geom_bar() 
```



