---
title: "Programming assignment 2"  
author: "A Valeria Carbone"  
date: "Last update: `r Sys.time()`"  
output:  
  html_document:  
    highlight: kate  
    keep_md: yes  
    theme: united---

```{r}
library (languageR)
library (tidyverse)
glimpse (danish)
glimpse (english)
glimpse (dativeSimplified)
```



# 1 Scatterplot
```{r}
beginningReaders |>
  ggplot() +
  aes(x = LogFrequency, y = ProportionOfErrors) +
  geom_point() +
  labs(x = "LogFrequency", y = "Proportion of Errors", title = "Log Frequency and Proportion of Errors") +
  theme_minimal()

```

# There does not seem to be a correlation between Log Frequency and the Proportion of Errors. There is a weak correlation because it looks scattered.



# 2. Boxplot

```{r}
danish |>
  ggplot() +
  aes(x = Affix, y = LogRT) +
  geom_boxplot(width = 0.2, alpha = 0.6) +
  labs(x = "Affix", y = "Log Reaction Time", title = "Reaction times per Affix") +
  theme_minimal()

```

# similar reaction times in means for all affixes, with some participants (outliers) taking much longer

# 3 Barplot
```{r}
english |>
  ggplot() +
  aes(x = AgeSubject, y = RTnaming, fill = AgeSubject)+
  geom_col (width = 0.2, alpha = 0.6) +
  labs(x = "Age Subject", y = "Reaction Time Naming", title = "Reaction times per Age Subject") +
  theme_minimal()

scale_fill_manual (values = c ("young" = "skyblue", "old" = "purple"))
```

# These two bars show that young and old reaction times for naming are almost identical.
















