---
title: "PA4"
author: "Stephanie Rodriguez"
institute: "Rutgers University"
date: "2023/03/27 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["hygge", "rutgers", "rutgers-fonts"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
# Load Packages
```{r, load-packages}
library(tidyverse)
library(ds4ling)
library(here)
library("broom")
```

---
# Load Data

```{r, load-data}
read_csv(file = "./data_raw/fricatives/fricatives.csv")
data_fricatives <- read_csv(file = "./data_raw/fricatives/fricatives.csv")

fricatives_data <- read.csv(here("data_raw","fricatives","fricatives.csv"))
```

---
# Tidy Data
```{r, data-tidy}
tidy_fricatives <- data_fricatives %>% 
  pivot_longer(
    cols = c("s_cog", "sh_cog"), 
    names_to = "phoneme", 
    values_to = "cog") %>% 
  pivot_longer(
    cols = c("s_skewness", "sh_skewness"), 
    names_to = "phoneme1", 
    values_to = "skewness") %>% 
  unite(phoneme , phoneme1) %>% 
  relocate(phoneme, .before = cog) 
```

---
# Create CSV

```{r, create-csv}
write.csv(tidy_fricatives, "./data_tidy/tidy_fricatives.csv")
```

---
# Descriptive Stats
```{r, descriptive-stats}
tidy_fricatives %>% 
  group_by(phoneme) %>% 
  summarise(mean_cog = mean(cog), sd_cog = sd(cog), mean_skewness = mean(skewness), sd_skewness = sd(skewness)) %>% 
  knitr::kable()
```

---
# Plot of COG
```{r, boxplot}
tidy_fricatives %>% 
ggplot(aes(x = phoneme, y = cog)) + 
  geom_boxplot() + 
  labs(title , "Center of Gravity as a Function of Phoneme",
       x = "Phoneme",
       y = "COG")
```

---
# Plot of Skewness
```{r, plot}
tidy_fricatives %>% 
  ggplot(aes(x = phoneme, y = skewness)) +
  geom_point() + 
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  labs(title , "Skewness as a Function of Phoneme",
       x = "Phoneme",
       y = "Skewness")
```

---
# Model Summary
```{r, model-summary}
phoneme_s <- filter(tidy_fricatives, phoneme == "s_skewness")

model_0 <- lm(cog ~ 1, data = phoneme_s)
model_s <- lm(cog ~ skewness, data = phoneme_s)

summary(model_s)
tidy(model_s) %>% 
  knitr::kable(format = "pandoc")
         
```

---
# Scatterplot
```{r}
tidy_fricatives %>%
  ggplot(aes(x = skewness, y = cog)) + 
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "Skewness", y = "COG", title = "Center of Gravity as a Function of Skewness for the [s]")
```

---
# Model Diagnostics Anova
```{r, anova-diagnostics}
anova(model_0, model_s)
```

---
# Assumptions
```{r, assumptions-data}
assumptions_data <- read.csv(here("data_raw","fricatives","assumptions.csv"))
```

---
# Assumptions Plot
```{r, assumptions-plot}
assumptions_data %>% 
  ggplot(aes(x = x, y = y )) +
  geom_point() 
```

---
# Results

A linear regression analysis was performed to examine the relationship between the center of gravity and the skewness of the phoneme [s]. The dependent variable was center of gravity and the predictor was and skewness. 

The results of the analysis indicated a positive linear relationship between the two variables, as evidenced by the plot. The adjusted R-squared value of 0.4024 indicated that 40.24% of the variation in the center of gravity could be explained by skewness. Furthermore, a one unit increase in skewness was associated with an increase of 2664.3 in the center of gravity, with a standard error of 526.7. The statistical significance of the model was confirmed by a p-value of 1.105e-05, indicating that the model as a whole was significant at a level of alpha = 0.05.

The model revealed a main effect of skewness on center of gravity.







